---
title: "Week 5 Code Along Penguin Size Comparisons"
author: "Carly Norris"
date: "11/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests
-Used to compare means between groups

Mass comparison: Male & female chinstrap penguins

Do exploring becuase we always should

```{r}
chinstraps <- penguins %>% 
  filter(species== "Chinstrap")

# Histograms- only takes one variable and makes the counts for us
ggplot(data= chinstraps, aes(x= body_mass_g))+
  geom_histogram(bins= 10)+
  facet_wrap(~sex)# will split two groups male and female into different plots

```
The distribution looks relatively normal male is slightly larger than female.


## Quantaile-quantile plot compaires our sample to a sample that is normally distributed
```{r}
ggplot(data= chinstraps, aes(sample= body_mass_g))+
  geom_qq()+
  facet_wrap(~sex)
```
These are as close to normally distriuted as you can hope your data will be. 

## Get descriptive statistics


```{r}
chinstraps %>% #dont need to save this outcome
  group_by(sex) %>% 
  summarize( # creates new columns 
    mean_mass = mean(body_mass_g), #labels the column names
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
```

Get vectors of body_mass_g for female and male chinstrap pemguins separately.

```{r}
chinstrap_f <- chinstraps %>% 
  filter(sex== "female") %>% 
  pull(body_mass_g) #will pull out the values just from the variables we specify

chinstrap_m <- chinstraps %>% 
    filter(sex== "male") %>% 
  pull(body_mass_g)
```
### T-tests, a couple of different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstrap_f, chinstrap_m)
chinstrap_ttest

cohen.d(chinstrap_f, chinstrap_m)
```
t= how different from zero
degrees of freedom= measure of sample size

d estimate is large which means there is a clear difference between female and male penguins. This isn't as highly impacted by sample sizes. Broom package tidy package. Look in code along key. 

2 Are male chinstrap penguins significantly larger (body mas) than female chinstrap penguins?

The order of things matters. 

```{r}
t.test(chinstrap_m,chinstrap_f, alternative = "greater") #instead of two sided test update function to make it onesided
```
## One-way ANOVA

NOVA way to compare means across more than two groups. ONE-WAY means we have a single factor that has discreat groups within it. 
### Make table of counts by species and sex

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)

penguin_counts
```
Create a subset of only female penguins

```{r}
penguin_f <- penguins %>% 
  filter(sex =="female")

ggplot(data=penguin_f, aes(x= flipper_length_mm))+
  geom_histogram(aes(fill= species))+
  facet_wrap(~species, scales = "free") #males the three different graphs and each pannel doesn't have to have the scale good for exploring distributions within groups

ggplot(data=penguin_f, aes(sample= flipper_length_mm))+
  geom_qq(aes(color= species))+
  facet_wrap(~species, scales = "free")
```

### Find summary statistics

starting from penguins_f, create a sum table that groups by species and returs the following

-mean of flipper length
-sd
-sample size
-variance
-standard error

```{r}
penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize( 
    mean_flip = mean(flipper_length_mm), 
    sd_flip = sd(flipper_length_mm),
    sample_size = n(),
    se_flip = sd(flipper_length_mm)/sqrt(n()),
    variance_flip = var(flipper_length_mm)
    )

leveneTest(flipper_length_mm ~ species, data=penguin_f) # in car package hypothesis test null hypothesis our goups have equal variances

```
Levene's large sample careful because could get significance

### Beeswarm plot with mean +- 1sd

```{r}
ggplot()+ 
  geom_beeswarm(data= penguin_f, aes(x=species, y=flipper_length_mm),
                color= "grey70",
                size= 3,
                alpha = 0.5)+
  geom_point(data=penguin_f_summary, aes(x=species, y= mean_flip),
             size=4)+
  geom_errorbar(data=penguin_f_summary,
                aes(x=species,
                    ymax= mean_flip + sd_flip,
                    ymin= mean_flip- sd_flip),
                width= 0.1)
```

### One-way ANOVA

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov)

TukeyHSD(penguin_aov)
```
aov funtion by default runs a type one ANOVA. Usually if you have unequal sample size etc want a type three ANOVA

There is a very negligible possibilitiy that these samples are from the same pop. Reject the null. This allows us to reject the null that all saples have pops with the same mean pop. Want to do a post hoc test that does pairwise tests. 

See that all are below .05 so there is a sig diff between all pairwise comparisons. 
If you have very unbalanced design that you are trying to do an ANOVA then you want to do an orthoginal or type three ANOVA

### Type III Anova with the car Package

```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f) #finds a linear relationship between flip length and species
penguin_aov_t3 <- Anova(penguin_lm, type= "III")
penguin_aov_t3
```











