---
title: "Week 5 Code Along Penguin Size Comparisons"
author: "Carly Norris"
date: "11/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests
-Used to compare means between groups

Mass comparison: Male & female chinstrap penguins

Do exploring becuase we always should

```{r}
chinstraps <- penguins %>% 
  filter(species== "Chinstrap")

# Histograms- only takes one variable and makes the counts for us
ggplot(data= chinstraps, aes(x= body_mass_g))+
  geom_histogram(bins= 10)+
  facet_wrap(~sex)# will split two groups male and female into different plots

```
The distribution looks relatively normal male is slightly larger than female.


## Quantaile-quantile plot compaires our sample to a sample that is normally distributed
```{r}
ggplot(data= chinstraps, aes(sample= body_mass_g))+
  geom_qq()+
  facet_wrap(~sex)
```
These are as close to normally distriuted as you can hope your data will be. 

## Get descriptive statistics


```{r}
chinstraps %>% #dont need to save this outcome
  group_by(sex) %>% 
  summarize( # creates new columns 
    mean_mass = mean(body_mass_g), #labels the column names
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
```




